### 2018-12-20 test.Rmd
close to figuring out the weird stuff...

---
title: "Untitled"
output:
  word_document: 
    reference_docx: "manuscript/manuscript_template.docx"
---

# Plain Data Frame

```{r plain}
x<-data.frame(unit = "µg/l", stringsAsFactors = FALSE)
knitr::kable(x)
```

# Tibble
```{r tibble}
x<-dplyr::tbl_df(data.frame(unit = "µg/l", stringsAsFactors = FALSE))
knitr::kable(x, format = "markdown")
```

# My stuff
```{r culprit}
source(here::here("R/functions.R"))
ww_all<-read_csv(here("data/ww_lake_trend_data.csv"))
ww_param_ranges <- ww_all %>%
  filter(param == "chla" |
           param == "temp" |
           param == "total_n" |
           param == "total_p") %>%
  mutate(param = factor(param,levels = c("temp", "total_n", "total_p", "chla"))) %>%
  group_by(param) %>%
  summarize(min = round(min(mn_measurement, na.rm = T), 2),
            percentile_25 = round(quantile(mn_measurement, probs = 0.25, na.rm = T), 2),
            mean = round(mean(mn_measurement, na.rm = T), 2),
            median = round(median(mn_measurement, na.rm = T), 2),
            percentile_75 = round(quantile(mn_measurement, probs = 0.75, na.rm = T), 2),
            max = round(max(mn_measurement, na.rm = T), 2),
            sd = round(sd(mn_measurement, na.rm = T), 2)) %>%
  ungroup() %>%
  mutate(units = c("celsius", "µg/l", "µg/l","µg/l")) %>%
  select(Parameter = param, Units = units, 
         "25th Percentile" = percentile_25,
         Mean = mean, Median = median, 
         "75th Percentile" = percentile_75,
         Max = max, "Std. Dev" = sd )
knitr::kable(ww_param_ranges, format = "markdown")
```