### 2019-11-05 nalms_2019_talk/increasing_chla_stable_nutrients_talk.Rmd
outline and draft of slide 1 done

---
title: "Increasing productivity amid stable nutrient regmines in Rhode Island Lakes and Reservoirs"
subtitle: "NALMS 2019, Burlington VT"
author: "Jeff Hollister, Q Kellogg, Betty Kreakie, Stephen Shivers, Elizabeth Herron, Linda Green, Bryan Milstead, and Art Gold"
date: "2019-11-14"
output:  
  powerpoint_presentation:
    reference_doc: increasing_chla_stable_nutrients_template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What drives lake productivity
::::::::::::::{.columns}
:::{.column}
- Nutrients
- Temperature
- Other things
:::
:::{.column}

![](figures/es-2016-02575v_0006.jpeg)

![](figures/tn_tp_chla_plot.jpg)
:::
::::::::::::::

## Nutrient Management

- One goal is to reduce increased phyto
- What has happened with P and N reductions?
- Did it work?
  - Long-term data needed

## Stasis in Northeast

content

## But change at other scales

- VT example
- Other examples?
- What about RI?

## Watershed Watch

- Details on WW program and data collected

## Watershed Watch: filtered for trend analysis

- What we filtered
- resultant datasets details

## LAGOSNE

## Analysis: Site-specific z-scores

## Analysis: Yearly trends

## Results: Nutrient trends

## Results: Chlorophyll trends

## Results: Temperature trends

## Conclusions

### In Rhode Island
- More chlorophyll
- More heat
- Nutrients not changing much

### Within the Northeast
- Stasis (our analysis concurs)

## Thank you and Questions