### 2019-11-07 nalms_2019_talk/increasing_chla_stable_nutrients_talk.Rmd
progress on slides!

---
title: "Increasing productivity amid stable nutrient regmines in Rhode Island Lakes and Reservoirs"
subtitle: "NALMS 2019, Burlington VT"
author: "Jeff Hollister, Q Kellogg, Betty Kreakie, Stephen Shivers, Elizabeth Herron, Linda Green, Bryan Milstead, and Art Gold"
date: "2019-11-14"
output:  
  powerpoint_presentation:
    reference_doc: increasing_chla_stable_nutrients_template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What drives changes in lake productivity?
::::::::::::::{.columns}
:::{.column}
- Nutrients
- Temperature
- Other things
:::
:::{.column}

![](figures/es-2016-02575v_0006.jpeg)

![](figures/tn_tp_chla_plot.jpg)

![](figures/temp_chla_plot.jpg)
:::
::::::::::::::

## The management problem

- Desire to reduce and manage nutrient pollution
- Phosphorus Bans
- Nutrient Criteria
- Enacted piecemeal over last ~20 years
- Do the data show any change?

## Stasis in Northeast

- Oliver et al (2017) Unexpected stasis in a changing world: Lake nutrient and
chlorophyll trends since 1990. Global Change Biology. DOI: 10.1111/gcb.13810

![](figures/10.1111-gcb.13810Figure138100001.jpg)

## But change at other scales

- Mathews et al (2018) Is Vermont losing its oligotrophic lakes?. LakeLine. 38-2.
- Rhode Island?

![](figures/mathews_vt_olgio_lakeline_fig2.jpg)

## URI Watershed Watch

- Volunteer monitoring program
  - ~400 volunteers
- Started in 1988 with 14 lakes
- Now
  - 250+ sites
  - 120+ waterbodies
- Rigorous QA/QC
- Data used by RI DEM and US EPA

## URI Watersehd Watch: Lake Sampling

- May to October
- Weekly
  - Water Clarity
  - Temperature
- Every Other Week
  - Chlorophyll
  - Dissolved Oxygen
- 3 Times per Season
  - Nutrients
  - Alakalinity
  - pH
  - Bacteria

## URI Watershed Watch: Filtered for Trend Analysis

- Parameters: Total Nitrogen, Total Phosphorus, N:P, Temperature, Chlorophyll
- Years: 1993 to 2016
- Months: May through October
- Depths: <= 2 meters
- Sites: At least 10 years of data
- Total of 69 Sites

![](figures/ww_map.jpg)

## LAGOSNE

## Analysis: Site-specific z-scores

## Analysis: Yearly trends

## Results: Nutrient trends

## Results: Chlorophyll trends

## Results: Temperature trends

## Conclusions

### In Rhode Island
- More chlorophyll
- More heat
- Nutrients not changing much

### Within the Northeast
- Stasis (our analysis concurs)

## Thank you and Questions