### 2019-11-08 nalms_2019_talk/increasing_chla_stable_nutrients_talk.Rmd

---
title: "Increasing productivity amid stable nutrient regmines in Rhode Island Lakes and Reservoirs"
subtitle: "NALMS 2019, Burlington VT"
author: "Jeff Hollister, Q Kellogg, Betty Kreakie, Stephen Shivers, Elizabeth Herron, Linda Green, Bryan Milstead, and Art Gold"
date: "2019-11-14"
output:  
  powerpoint_presentation:
    reference_doc: increasing_chla_stable_nutrients_template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What drives changes in lake productivity?
::::::::::::::{.columns}
:::{.column}
- Nutrients
- Temperature
- Other things
:::
:::{.column}

![](figures/es-2016-02575v_0006.jpeg)

![](figures/tn_tp_chla_plot.jpg)

![](figures/temp_chla_plot.jpg)
:::
::::::::::::::

## The management problem

- Desire to reduce and manage nutrient pollution
- Phosphorus Bans
- Nutrient Criteria
- Enacted piecemeal over last ~20 years
- Do the data show any change?

## Stasis in Northeast

- Oliver et al (2017) Unexpected stasis in a changing world: Lake nutrient and
chlorophyll trends since 1990. Global Change Biology. DOI: 10.1111/gcb.13810

![](figures/10.1111-gcb.13810Figure138100001.jpg)

## But change at other scales

- Mathews et al (2018) Is Vermont losing its oligotrophic lakes?. LakeLine. 38-2.
- Rhode Island?

![](figures/mathews_vt_olgio_lakeline_fig2.jpg)

## URI Watershed Watch

- Volunteer monitoring program
  - ~400 volunteers
- Started in 1988 with 14 lakes
- Now
  - 250+ sites
  - 120+ waterbodies
- Rigorous QA/QC
- Data used by RI DEM and US EPA

## URI Watersehd Watch: Lake Sampling

- May to October
- Weekly
  - Water Clarity
  - Temperature
- Every Other Week
  - Chlorophyll
  - Dissolved Oxygen
- 3 Times per Season
  - Nutrients
  - Alakalinity
  - pH
  - Bacteria

## URI Watershed Watch: Filtered for Trend Analysis

- Parameters: Total Nitrogen, Total Phosphorus, N:P, Temperature, Chlorophyll
- Years: 1993 to 2016
- Months: May through October
- Depths: <= 2 meters
- Sites: At least 10 years of data
- Total of 69 Sites
  - 66 with all parameters
  - 1 without N:P
  - 2 without chlorophyll and temperature

![](figures/ww_map.jpg)

## LAke mult-scaled GeOSpatial and temoporal database for the Northeastern US (LAGOS-NE)

- Soranno et al (2017). LAGOS-NE: a multi-scaled geospatial and temporal database of lake ecological context and water quality for thousands of US lakes. Gigascience, 6(12)
- Stachelek and Oliver (2017). LAGOSNE: Interface to the lake multi-scaled geospatial and temporal database. R package version 1.1.0. https://cran.r-project.org/package=LAGOSNE
- <https://lagoslakes.org/>
- 17 States: CT, IL, IN, IA, ME, MA, MI, MN, MO, NH, NJ, NY, OH, PA, RI, VT, WI
- ~ 50,000 Lakes

![](figures/gix101fig1.jpeg)

## LAke mult-scaled GeOSpatial and temoporal database for the Northeastern US (LAGOS-NE): Filtered for Trend Analysis

- Parameters: Total Nitrogen, Total Phosphorus, N:P, Chlorophyll
  - Temperature data not currently available
- Years: 1993 to 2016
- Months: May through October
- Depths: <= 2 meters
- Sites: At least 10 years of data
- Total of 1482 Sites
  - 637 with chlorophyll and phosphorus
  - 377 with only chlorophyll
  - 275 with only phosphorus
  - 161 with all variables (except temperature)
  - 25 with chlorophyll and nitrogen
  - 7 with other combination of nutrients

## Analysis: Site-specific z-scores

![](figures/z-score_schematic.jpg)

## Analysis: Yearly trends

## Results: Nitrogen trends

![](figures/ww_tn_trends.jpg)

![](figures/lagos_tn_trends.jpg)

## Results: Phosphorus trends

![](figures/ww_tp_trends.jpg)

![](figures/lagos_tp_trends.jpg)

## Results: N:P trends

![](figures/ww_np_trends.jpg)

![](figures/lagos_np_trends.jpg)

## Results: Chlorophyll trends

![](figures/ww_chla_trends.jpg)

![](figures/lagos_chla_trends.jpg)

## Results: Temperature trends

![](figures/ww_temp_trends.jpg)

## Conclusions

### In Rhode Island
- More chlorophyll
- More heat
- Nutrients not changing much

### Within the Northeast
- Stasis (our analysis concurs)

## Thank you and Questions