### 2020-09-21 table.Rmd

---
title: "Untitled"
author: "Jeff Hollister"
date: "9/15/2020"
output:
  word_document:
    reference_docx: manuscript/manuscript_template.docx
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pander)
library(knitr)
library(kableExtra)
library(dplyr)
```

## Tables

```{r table, echo=FALSE}
mydf<-tibble(`Big Long Column Name Number1` = rnorm(10,1000,100),
       `Big Long Column Name Number2` = rnorm(10,1000,100),
       `Big Long Column Name Number3` = rnorm(10,1000,100),
       `Big Long Column Name Number4` = rnorm(10,1000,100),
       `Big Long Column Name Number5` = rnorm(10,1000,100),
       `Big Long Column Name Number6` = rnorm(10,1000,100)
       )
```

```{r kable, results="asis"}
kable(mydf,format = "markdown")
```